The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-05-02 writeup.tex
added all confidence intervals and editted some more

\begin{figure}

{\centering \includegraphics{writeup_files/figure-latex/linkvisualization-1}

}

\caption{Strong and weak link from response options to researcher inference about scalar implicature rate, exemplified for the disjunctive utterance when the conjunction is true.}\label{fig:linkvisualization}
\end{figure}

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/stimuli-1}

}

\caption{Cards used in the connective guessing game.}\label{fig:stimuli}
\end{figure}

@@

Participants could derive implicatures in two trial types. First, the
trial type in which two animals were present on the card (e.g.~cat and
dog) but Bob guessed only one of them (e.g. \enquote{there is a cat}).
Such trials can have a literal interpretation (cat and possibly more) or
an exhaustive interpretation (only cat). We refer to them as
\enquote{exhaustive}. The second trial type with implicatures was the
one in which two animals were on the card (e.g.~cat and dog) and Bob
used a disjunciton (e.g.~cat or dog). These trials can have a literal
(inclusive) interpretation (e.g.~cat or dog or both), or an exclusive
interpretation (e.g.~cat or dog, not both). We refer to these trials as
\enquote{scalar}. The following four trial types were used as
experimental control: two trial types where there was no overlap between
the guess (e.g.~elephant) and the animal(s) on the card (e.g.~cat, cat
and dog); and two trial types where the animal(s) on the card were
correctly guessed. For example, if there was only a cat on the card, Bob
said \enquote{there is a cat} and if there was a cat and a dog, Bob said
\enquote{there is a cat and a dog}.

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/trials-1}

}

\caption{Trial types represented by example cards and guesses.}\label{fig:trials}
\end{figure}

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/binaryPlot-1}

}

\caption{Proportion responses for the two-alternative (binary) forced choice judgments in the guessing game.}\label{fig:binaryPlot}
\end{figure}

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/ternaryPlot-1}

}

\caption{Proportion responses for the three-alternative (ternary) forced choice judgments in the guessing game.}\label{fig:ternaryPlot}
\end{figure}

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/quaternaryPlot-1}

}

\caption{Proportion responses for the four-alternative (quatenary) forced choice judgments in the guessing game.}\label{fig:quaternaryPlot}
\end{figure}

@@

\begin{figure}[t]

{\centering \includegraphics{writeup_files/figure-latex/quinaryPlot-1}

}

\caption{Proportion responses for the five-alternative (quinary) forced choice judgments in the guessing game.}\label{fig:quinaryPlot}
\end{figure}

@@

\subsubsection{Analysis}\label{analysis}

The goal of this study is to see if the estimated \enquote{implicature
rate} in the experimental task is affected by the linking assumptions
and the number of response options available in the task. Our analysis
in this section focuses on these three elements. Two trial types are
predicted to include pragmatic implicatures. First, trials where two
animals were on the card but the fictional character guessed with a
disjunction; for example \enquote{cat or dog} when the card has both a
cat and a dog on it. We called such trials \enquote{scalar}. Second,
trials where there were two animals on the card but the character
guessed only one; for example \enquote{cat} when the card had a cat and
a dog on it. We called such trials \enquote{exhaustive}. In our
assessment of implicature rate, we focus on these two types of trials.

This study set out to test two hypotheses. First, that the proportion of
pragmatic vs.~literal responses in a truth values judgement task changes
based on the number of response options available to the participants.
We test this hypothesis formally using a binomial mixed effects model
with the fixed effect of response type and the random intercept for
participants as well as random intercept and slope for

A second hypothesis was that the definition of what responses count as
participants computing an implicature may affect the estimated rate of
implicature in the experimental task.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/implicatureRatePlot-1.pdf}
\caption{}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/linkvisualization-1.pdf}
\caption{\label{fig:linkvisualization}Strong and weak link from response
options to researcher inference about scalar implicature rate,
exemplified for the disjunctive utterance when the conjunction is true.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/stimuli-1.pdf}
\caption{\label{fig:stimuli}Cards used in the connective guessing game.}
\end{figure}

@@

Participants could derive implicatures in two trial types. First, the
trial type in which two animals were present on the card (e.g.~cat and
dog) but Bob guessed only one of them (e.g. \enquote{there is a cat}).
Such trials can have a literal interpretation (cat and possibly more) or
an exhaustive interpretation (only cat). We refer to them as
\enquote{exhaustive}. The second trial type with implicatures was the
one in which two animals were on the card (e.g.~cat and dog) and Bob
used a disjunciton (e.g.~cat or dog). These trials can have a literal
(inclusive) interpretation (e.g.~cat or dog or both), or an exclusive
interpretation (e.g.~cat or dog, not both). We refer to these trials as
\enquote{scalar}. The following four trial types were used as
experimental control: two trial types where there was no overlap between
the guess (e.g.~elephant) and the animal(s) on the card (e.g.~cat, cat
and dog); and two trial types where the animal(s) on the card were
correctly guessed. For example, if there was only a cat on the card, Bob
said \enquote{there is a cat} and if there was a cat and a dog, Bob said
\enquote{there is a cat and a dog}. Since the fictional character was
blindfolded and did not see the outcome of the game, the ignorance
inference commonly associated with disjunction was already common ground
in the experimental setting. If the character was seeing the cards or
knew what was on them, a disjunction would have violated the expectation
that the speaker does not know which alternative actually holds. Our
study controls for the possibile effect of ignorance violations on
exclusivity and exhaustive inferences.

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/trials-1.pdf}
\caption{\label{fig:trials}Trial types represented by example cards and
guesses.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/binaryPlot-1.pdf}
\caption{\label{fig:binaryPlot}Proportion responses for the two-alternative
(binary) forced choice judgments in the guessing game.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/ternaryPlot-1.pdf}
\caption{\label{fig:ternaryPlot}Proportion responses for the
three-alternative (ternary) forced choice judgments in the guessing
game.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/quaternaryPlot-1.pdf}
\caption{\label{fig:quaternaryPlot}Proportion responses for the
four-alternative (quatenary) forced choice judgments in the guessing
game.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/quinaryPlot-1.pdf}
\caption{\label{fig:quinaryPlot}Proportion responses for the
five-alternative (quinary) forced choice judgments in the guessing
game.}
\end{figure}

@@

\subsection{Analysis}\label{analysis}

Our primary goal in this study was to check whether the estimated
\enquote{implicature rate} in the experimental task is affected by the
linking assumption and the number of response options available in the
task. Our analysis in this section focuses on these three elements. As
mentioned before, two trial types were predicted to include pragmatic
implicatures. First, trials where two animals were on the card but the
fictional character guessed with a disjunction (scalar); for example
\enquote{cat or dog} when the card has both a cat and a dog on it.
Second, trials where there were two animals on the card but the
character guessed only one (exhaustive); for example \enquote{cat} when
the card had a cat and a dog on it. We called such trials
\enquote{exhaustive}. In our assessment of implicature rate, we focus on
these two trial types.

We considered two linking assumptions. First we defined a weak (lenient)
linking assumption in which any response lower than the maximum point on
the scale (i.e. \enquote{right}) is considered evidence for implicature
computation. Second, we defined a strong (strict) linking assumption
that only considered the lowest point on the scale (i.e.
\enquote{wrong}) as evidence for implicature computation. For each
condition in our study (binary, ternary, quaternary, and quinary) and
each implicature trial type (exhaustive and scalar), we computed a weak
and a strong implicature rate. Figure \ref{fig:implicatureRatePlot}
shows these computed rates.

Comparing the strong and weak rows on Figure
\ref{fig:implicatureRatePlot}, we see that a weak linking assumption
tends to estimate higher implicature rates, especially in tasks with
more response options. With a strong linking assumption, the binary and
possibly ternary judgment tasks derive higher implicature rates than
quaternary and quinary tasks. With a weak linking assumption, the
pattern is reversed. Quaternary and quinary tasks estimate higher rates
than binary and ternary tasks.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/implicatureRatePlot-1.pdf}
\caption{\label{fig:implicatureRatePlot}Implicature rate in exhaustive and
scalar trials of the binary, ternary, quaternary, and quinary versions
of the guessing game, computed once with a strong linking assumption and
once with a weak linking assumption.}
\end{figure}

We test this hypothesis formally using a binomial mixed effects model
with the fixed effect of response type and the random intercept for
participants as well as random intercept and slope for

All the analytical decisions described here were pre-registered and you
can access them through the following link:
\url{https://aspredicted.org/tq3sz.pdf}
